### 2024-03-28 paper/paper.qmd
updated test_data and paper

---
title: "Analysis of Auto Thefts in Ottawa, Canada"
subtitle: "My subtitle if needed"
author: 
  - Catherine Punnoose
thanks: "Code and data are available at: https://github.com/catherinee1216/Auto_Theft_Crime_Analysis_Ottawa.git."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-03-28 scripts/02-test_data.R

@@ -11,6 +11,7 @@
 #### Workspace setup ####
 library(arrow)
 library(tidyverse)
-# [...UPDATE THIS...]
+
+# use testthat package
 
 #### Test data ####

### 2024-03-28 paper/paper.qmd
more updates

@@
   - Catherine Punnoose
-thanks: "Code and data are available at: https://github.com/catherinee1216/Auto_Theft_Crime_Analysis_Ottawa.git."
+thanks: "Code and data are available at: https://github.com/catherinee1216/AutoTheft__Ottawa.git."
 date: today
@@
 # Results
+```{r}
+
+things i can add 
+# averages 
+# differences 
+# summary stats 
+# changes between point A and point A
+
+```
 

### 2024-03-28 scripts/01-download_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from [...UPDATE THIS...]
 # Author: Rohan Alexander [...UPDATE THIS...]
 # Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
@@ -9,11 +9,18 @@
 
 
 #### Workspace setup ####
+library(opendatatoronto)
 library(tidyverse)
 # [...UPDATE THIS...]
 
-#### Load data ####
-# [...ADD CODE HERE...]
+#### Download data ####
+# [...ADD CODE HERE TO DOWNLOAD...]
 
 
 
+#### Save data ####
+# [...UPDATE THIS...]
+# change the_raw_data to whatever name you assigned when you downloaded it.
+write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+
+         

### 2024-04-02 scripts/01-download_data.R
updated download_data, data_cleaning, paper

@@ -1,26 +1,25 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data 
+# Author: Catherine Punnoose
+# Date: 03 April 2024
+# Contact: catherine.punnoose@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Auto_Theft_ottawa_Data.csv 
 
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
+# code from CHAT GPT
+# Set the file path to your CSV file
 
+# Read the CSV file
+auto_theft_data <- read.csv("data/analysis_data/Auto_Theft_Ottawa_Data.csv")
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+# saving original dataset in raw_data
+write.csv(auto_theft_data, "data/raw_data/Auto_Theft_Ottawa_Data.csv") 
 
          

### 2024-04-02 scripts/02-data_cleaning.R

@@ -8,38 +8,24 @@
 # Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
+library(dplyr)
+library(tidyr)
+library(janitor)
 library(tidyverse)
 library(arrow)
 
 #### Clean data ####
-raw_data <- read_csv("data/analysis_data/Auto_Theft_Ottawa_Data.csv")
+raw_data <- read.csv("data/raw_data/Auto_Theft_Ottawa_Data.csv")
 
 cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+  select(vehicle_make, vehicle_model, vehicle_year, vehicle_colour, 
+         vehicle_value, occurred_date, reported_date, weekday, 
+         recovered, year) |>
+  mutate(theft_year = year) # changing year to theft_year for clarity
+
 
 #### Save data ####
-write_parquet(cleaned_data, "data/analysis_data/Auto_Theft_Ottawa_Data.csv")
+write.csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+

### 2024-04-03 paper/paper.qmd
updated download_data, data_cleaning, paper
updated paper

--- 2024-03-28 paper/paper.qmd
@@
 ---
-title: "Analysis of Auto Thefts in Ottawa, Canada"
-subtitle: "My subtitle if needed"
+title: "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
 author: 
   - Catherine Punnoose
-thanks: "Code and data are available at: https://github.com/catherinee1216/Auto_Theft_Crime_Analysis_Ottawa.git."
+thanks: "Code and data are available at: https://github.com/catherinee1216/AutoTheft_Ottawa.git."
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The surge in auto theft rates across Canada has prompted heightened vigilance among car owners regarding their vehicle's security.
+This paper delves into a thorough examination of auto theft occurrences within Canada's capital, Ottawa.
+Focusing on the years 2018-2022, the paper provides a deep dive into the types of vehicles brands, model, and color targeted and offers insights into the neighborhoods within Ottawa exhibiting elevated rates of auto theft.
+Additionally, the study assesses factors such as the recovery rate of stolen cars, and provides solutions to improve vehicle security by understanding auto theft dynamics in the region."
 format: pdf
@@
 library(tidyverse)
-library(palmerpenguins)
+library(janitor)
+library(kableExtra)
+library(maps)
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+In recent years, there is a significant increase of auto theft rates across various cities in Canada.
+The country’s capital, Ottawa has reported 221 stolen vehicles cases since January 1, 2024 [@citeCTV].
+This escalating trend in auto theft poses a considerable risk to public safety.
+Although Ottawa police have made efforts to mitigate such reports as well as educate the public, the number of reported cases continues to rise steadily, specifically from the years 2018-2023.
+However, questions regarding potential patterns in the types of cars targeted and the proportion of vehicles successfully recovered following theft have increased.
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper aims to identify the gravity of auto theft as a pressing concern, particularly in Canada, in the @Introduction.
+Subsequently, the @Data section focuses on the methodology for data collection and data cleaning.
+The @Results section delves into a comprehensive analysis of auto theft patterns in Ottawa from the years 2018-2022.
+Finally, in the @Discussion section, the paper explores potential drivers behind the escalation in auto theft rates, strategies for enhancing vehicle security, and prospective trends in auto theft.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
+#| include: false
 #| warning: false
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+auto_Data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
-
-Talk way more about it.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+```{r}
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+auto_Data |>
+  ggplot(aes(x = theft_year)) +
+  geom_histogram(stat = "count")
 ```
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
 ## Second discussion point
@@
 ## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 
@@
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-04-03 paper/paper.qmd
updated paper

@@
 date-format: long
-abstract: "The surge in auto theft rates across Canada has prompted heightened vigilance among car owners regarding their vehicle's security.
-This paper delves into a thorough examination of auto theft occurrences within Canada's capital, Ottawa.
+abstract: "The surge in auto theft rates across Canada has prompted car owners to become more vigilant regarding their vehicle's security.
+This paper gives a thorough examination of auto theft occurrences within Canada's capital, Ottawa.
 Focusing on the years 2018-2022, the paper provides a deep dive into the types of vehicles brands, model, and color targeted and offers insights into the neighborhoods within Ottawa exhibiting elevated rates of auto theft.

### 2024-04-03 scripts/02-data_cleaning.R
update data cleaning

@@ -27,5 +27,5 @@ cleaned_data <-
 
 
 #### Save data ####
-write.csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+write_parquet(cleaned_data, "data/analysis_data/analysis_data.parquet")
 

### 2024-04-04 paper/paper.qmd
updated paper, data cleaning

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(maps)
+library(ggmap)
+library(arrow)
+
+auto_Data <- read_parquet(here::here("data/analysis_data/analysis_data.parquet"))
 ```
@@
 
+[TO BE EDITED]
+
 This paper aims to identify the gravity of auto theft as a pressing concern, particularly in Canada, in the @Introduction.
@@
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-auto_Data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
-
-```
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The data was cleaned by using the open source statistical programming language R [@citeR]...[TO UPDATE]
 
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: theft-year
+#| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
 #| echo: false
+#| warning: false
+#| message: false
 
@@
 ```
+```{r}
+#| label: theft-brand
+#| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
+#| echo: false
+#| warning: false
+#| message: false
+
+brand_theft_2018 <- auto_Data |>
+  filter(theft_year == "2018")
+
+brand_theft_2018 |>
+  ggplot(aes(x = vehicle_make, fill = vehicle_make)) +
+  geom_histogram(stat = "count")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# creating map from section 5.4.1 in the textbook
+
+# Define the bounding box for Ottawa
+bbox <- c(left = -75.9, bottom = 45.3, right = -75.5, top = 45.5)
+
+# Get the map of Ottawa using get_stadiamap function
+ottawa_stadia_map <- get_stadiamap(bbox, zoom = 11, maptype = "stamen_toner_lite")
+
+# Plot the map using ggmap
+ggmap(ottawa_stadia_map) # map prints
+
+# ---- not working -----------
+#filtering dataset to get auto theft reports per neighborhood in 2018
+
+#neighborhood_theft_2018 <- auto_Data |>
+#  filter(theft_year == "2018") |>
+#  select(neighbourhood, x, y)
+
+#print(neighborhood_theft_2018)
+
+# map does not print with elements plotted
+
+#ggmap(ottawa_stadia_map, extent = "normal", maprange = FALSE) +
+#  geom_point(data = neighborhood_theft_2018,
+#             aes(x = 45.3, y = 45.5, colour = neighbourhood),
+#             alpha = 0.7) +
+#  scale_color_brewer(name = "2019 Division", palette = "Set1") +
+#  labs(x = "Longitude",
+#       y = "Latitude") +
+#  theme_minimal() +
+#  theme(panel.grid.major = element_blank(),
+#        panel.grid.minor = element_blank())
+
+```
 
@@
 
-## First discussion point {#sec-first-point}
+## Possible Reasons for Increasing Auto Theft Rates
 
-## Second discussion point
+## Methods to Increase Vehicle Security
 
-## Third discussion point
+## Auto Theft Rates Projection
 

### 2024-04-04 scripts/02-data_cleaning.R

@@ -22,7 +22,7 @@ cleaned_data <-
   janitor::clean_names() |>
   select(vehicle_make, vehicle_model, vehicle_year, vehicle_colour, 
          vehicle_value, occurred_date, reported_date, weekday, 
-         recovered, year) |>
+         recovered, year, neighbourhood, x, y) |>
   mutate(theft_year = year) # changing year to theft_year for clarity
 
 

### 2024-04-22 paper/paper.qmd
updated graphs

@@
 ```{r}
-#| label: theft-year
+#| label: averageTheftYear
 #| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
@@
   ggplot(aes(x = theft_year)) + 
+  labs(x = "Year", y = "Count") + theme_minimal() +
   geom_histogram(stat = "count")
+
 ```
 ```{r}
-#| label: theft-brand
+#| label: theft-brands
 #| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
+#| fig-subcap: ["Illnesses","Visits to the doctor"]
+#| layout-ncol: 3
 #| echo: false
@@
 
+# Filter for Number of Cars Stolen by Brand in 2018 
 brand_theft_2018 <- auto_Data |>
@@
 
-brand_theft_2018 |>
-  ggplot(aes(x = vehicle_make, fill = vehicle_make)) +
-  geom_histogram(stat = "count")
+# filter for Top 10 Brands 
+brand_theft_2018_table <- brand_theft_2018 |>
+                          count(vehicle_make)
+
+brands2018_top10 <- brand_theft_2018_table |>
+  filter(n > 26, n != 95) |> 
+  mutate(vehicle_make = recode(vehicle_make,
+                             "CHEV" = "Chevrolet",
+                             "DODG" = "Dodge",
+                             "FORD" = "Ford",
+                             "GMC" = "GMC",
+                             "HOND" = "Honda",
+                             "HYUN" = "Hyundai",
+                             "LEXS" = "Lexus",
+                             "MAZD" = "Mazda",
+                             "NISS" = "Nissan",
+                             "TOYT" = "Toyota"
+                             ))
+brands2018_top10 |>
+  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2018")
+
+# Filter for Number of Cars Stolen by Brand in 2019 
+brand_theft_2019 <- auto_Data |>
+  filter(theft_year == "2019")
+
+# filter for Top 10 Brands 
+brandtheft2019_top10 <- brand_theft_2019 |>
+                          count(vehicle_make) |>
+                          filter(n > 25, n != 118) |> 
+                          mutate(vehicle_make = recode(vehicle_make,
+                                                        "CHEV" = "Chevrolet",
+                                                        "DODG" = "Dodge",
+                                                        "FORD" = "Ford",
+                                                        "GMC" = "GMC",
+                                                        "HOND" = "Honda",
+                                                        "HYUN" = "Hyundai",
+                                                        "LEXS" = "Lexus",
+                                                        "MAZD" = "Mazda",
+                                                        "NISS" = "Nissan",
+                                                        "TOYT" = "Toyota"
+                                                     ))
+brandtheft2019_top10 |>
+  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2019")
+
+# Filter for Number of Cars Stolen by Brand in 2020
+brand_theft_2020 <- auto_Data |>
+  filter(theft_year == "2020")
+
+# filter for Top 10 Brands 
+brandtheft2020_top10 <- brand_theft_2020 |>
+                          count(vehicle_make) |>
+                          filter(n > 21, n != 95) |> 
+                          mutate(vehicle_make = recode(vehicle_make,
+                                                        "CHEV" = "Chevrolet",
+                                                        "DODG" = "Dodge",
+                                                        "FORD" = "Ford",
+                                                        "GMC" = "GMC",
+                                                        "HOND" = "Honda",
+                                                        "HYUN" = "Hyundai",
+                                                        "LEXS" = "Lexus",
+                                                        "MERZ" = "Mercedes-Benz ",
+                                                        "MAZD" = "Mazda",
+                                                        "NISS" = "Nissan",
+                                                        "TOYT" = "Toyota"
+                                                     ))
+brandtheft2020_top10 |>
+  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2020")
+
+# Filter for Number of Cars Stolen by Brand in 2021 
+brand_theft_2021 <- auto_Data |>
+  filter(theft_year == "2021")
+
+# filter for Top 10 Brands 
+brandtheft2021_top10 <- brand_theft_2021 |>
+                          count(vehicle_make) |>
+                          filter(n > 21, n != 115) |> 
+                          mutate(vehicle_make = recode(vehicle_make,
+                                                        "CHEV" = "Chevrolet",
+                                                        "DODG" = "Dodge",
+                                                        "FORD" = "Ford",
+                                                        "GMC" = "GMC",
+                                                        "HOND" = "Honda",
+                                                        "HYUN" = "Hyundai",
+                                                        "LEXS" = "Lexus",
+                                                        "MERZ" = "Mercedes-Benz ",
+                                                        "MAZD" = "Mazda",
+                                                        "NISS" = "Nissan",
+                                                        "TOYT" = "Toyota"
+                                                     ))
+brandtheft2021_top10 |>
+  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2021")
+
+# Filter for Number of Cars Stolen by Brand in 2022
+brand_theft_2022 <- auto_Data |>
+  filter(theft_year == "2022")
+
+# filter for Top 10 Brands 
+brandtheft2022_top10 <- brand_theft_2022 |>
+                          count(vehicle_make) |>
+                          filter(n > 34, n != 124) |> 
+                          mutate(vehicle_make = recode(vehicle_make,
+                                                        "CHEV" = "Chevrolet",
+                                                        "DODG" = "Dodge",
+                                                        "FORD" = "Ford",
+                                                        "GMC" = "GMC",
+                                                        "HOND" = "Honda",
+                                                        "HYUN" = "Hyundai",
+                                                        "LEXS" = "Lexus",
+                                                        "MERZ" = "Mercedes-Benz ",
+                                                        "MAZD" = "Mazda",
+                                                        "NISS" = "Nissan",
+                                                        "TOYT" = "Toyota"
+                                                     ))
+brandtheft2022_top10 |>
+  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")
+
+# Filter for Number of Cars Stolen by Brand in 2023 
+brand_theft_2023 <- auto_Data |>
+  filter(theft_year == "2023")
+
+# filter for Top 10 Brands 
+brandtheft2023_top10 <- brand_theft_2023 |>
+                          count(vehicle_make) |>
+                          filter(n > 16, n != 36) |> 
+                          mutate(vehicle_make = recode(vehicle_make,
+                                                       "ACUR" = "Acura",
+                                                        "CHEV" = "Chevrolet",
+                                                        "DODG" = "Dodge",
+                                                        "FORD" = "Ford",
+                                                        "GMC" = "GMC",
+                                                        "HOND" = "Honda",
+                                                        "HYUN" = "Hyundai",
+                                                        "LEXS" = "Lexus",
+                                                        "MERZ" = "Mercedes-Benz ",
+                                                        "MAZD" = "Mazda",
+                                                        "NISS" = "Nissan",
+                                                        "TOYT" = "Toyota"
+                                                     ))
+brandtheft2023_top10 |>
+  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2023")
 
@@
 ```{r}
+#| label: theft-recovery
+#| fig-cap: "Recovery Rates from 2018 - 2023"
 #| echo: false
@@
 
-# creating map from section 5.4.1 in the textbook
+theft_recovery <- auto_Data |>
+  select(theft_year, recovered)
 
-# Define the bounding box for Ottawa
-bbox <- c(left = -75.9, bottom = 45.3, right = -75.5, top = 45.5)
+theft_recovery |>
+  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
+  geom_bar(position = "dodge2") +
+  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
+  theme(legend.position = "bottom")
+```
+```{r}
+#| label: theft-neighborhoods
+#| fig-cap: "Recovery Rates from 2018 - 2023"
+#| echo: false
+#| warning: false
+#| message: false
 
-# Get the map of Ottawa using get_stadiamap function
-ottawa_stadia_map <- get_stadiamap(bbox, zoom = 11, maptype = "stamen_toner_lite")
+theft_recovery <- auto_Data |>
+  select(neighbourhood) |> 
+  count(neighbourhood) |> 
+  filter(n > 180)
 
-# Plot the map using ggmap
-ggmap(ottawa_stadia_map) # map prints
-
-# ---- not working -----------
-#filtering dataset to get auto theft reports per neighborhood in 2018
-
-#neighborhood_theft_2018 <- auto_Data |>
-#  filter(theft_year == "2018") |>
-#  select(neighbourhood, x, y)
-
-#print(neighborhood_theft_2018)
-
-# map does not print with elements plotted
-
-#ggmap(ottawa_stadia_map, extent = "normal", maprange = FALSE) +
-#  geom_point(data = neighborhood_theft_2018,
-#             aes(x = 45.3, y = 45.5, colour = neighbourhood),
-#             alpha = 0.7) +
-#  scale_color_brewer(name = "2019 Division", palette = "Set1") +
-#  labs(x = "Longitude",
-#       y = "Latitude") +
-#  theme_minimal() +
-#  theme(panel.grid.major = element_blank(),
-#        panel.grid.minor = element_blank())
-
+theft_recovery |>
+  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022") 
 ```
@@
 
-# Model details {#sec-model-details}
-
 \newpage

### 2024-04-22 scripts/02-data_cleaning.R

@@ -22,7 +22,7 @@ cleaned_data <-
   janitor::clean_names() |>
   select(vehicle_make, vehicle_model, vehicle_year, vehicle_colour, 
          vehicle_value, occurred_date, reported_date, weekday, 
-         recovered, year, neighbourhood, x, y) |>
+         recovered, year, neighbourhood) |>
   mutate(theft_year = year) # changing year to theft_year for clarity
 
 

### 2024-04-22 paper/paper.qmd
more graphs

@@
 
+auto_Data |>
+  ggplot(aes(x = theft_year)) + 
+  labs(x = "Year", y = "Count") + theme_minimal() +
+  geom_histogram(stat = "count")
+
 theft_recovery <- auto_Data |>
@@
   theme(legend.position = "bottom")
+
+# Merging 'theft_recovery' dataframe with 'auto_Data'
+merged_data <- merge(auto_Data, theft_recovery, by = "theft_year", all.x = TRUE)
+
+# Plotting the merged data as a line graph
+ggplot(merged_data, aes(x = theft_year, y = recovered)) +
+  geom_line() + 
+  geom_point() +
+  labs(x = "Year", y = "Count") +
+  theme_minimal()
 ```
@@
 
+years_vehicle <- auto_Data |>
+  select(vehicle_year) |> 
+  count(vehicle_year)
+
 theft_recovery |>
@@
   labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")
+
+library(ggplot2)
+
+# Assuming 'years_vehicle' is your data frame
+
+years_vehicle <- years_vehicle |>
+  mutate(vehicle_year = as.numeric(vehicle_year))  # Convert to numeric if 'vehicle_year' is factor
+
+# Create a scatter plot
+ggplot(years_vehicle, aes(x = vehicle_year, y = n, fill = n)) +
+  geom_point() +  # Use geom_point() for scatter plot
+  scale_x_continuous(limits = c(1930, 2024)) +  # Adjust the x-axis limits here
+  theme_minimal() + 
+  theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")
+
+```
+
+```{r}
+#| label: theft-neighborhoods
+#| fig-cap: "Recovery Rates from 2018 - 2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+theft_recovery <- auto_Data |>
+  select(neighbourhood) |> 
+  count(neighbourhood)
+
+library(ggplot2)
+
+# Assuming 'years_vehicle' is your data frame
+
+years_vehicle <- theft_recovery |>
+  mutate(vehicle_year = as.numeric(neighbourhood))  # Convert to numeric if 'vehicle_year' is factor
+
+# Create a scatter plot
+ggplot(years_vehicle, aes(x = neighbourhood, y = n, fill = neighbourhood)) +
+  geom_point() +  
+  theme_minimal() + 
+  theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")
+
 ```

### 2024-04-22 paper/paper.qmd
updated paper

@@
 
-auto_Data |>
-  ggplot(aes(x = theft_year)) + 
-  labs(x = "Year", y = "Count") + theme_minimal() +
-  geom_histogram(stat = "count")
+library(ggplot2)
 
@@
   select(theft_year, recovered)
+
+theft_count <- auto_Data |>
+  select(theft_year)
+
+theft_count |>
+  ggplot(mapping = aes(x = theft_year)) +
+  geom_bar() +
+  labs(x = "Age group", y = "Number of observations") +
+  theme(legend.position = "bottom")
 
@@
 
-# Merging 'theft_recovery' dataframe with 'auto_Data'
-merged_data <- merge(auto_Data, theft_recovery, by = "theft_year", all.x = TRUE)
+```
 
-# Plotting the merged data as a line graph
-ggplot(merged_data, aes(x = theft_year, y = recovered)) +
-  geom_line() + 
-  geom_point() +
-  labs(x = "Year", y = "Count") +
-  theme_minimal()
-```
 ```{r}
@@
 
-```{r}
-#| label: theft-neighborhoods
-#| fig-cap: "Recovery Rates from 2018 - 2023"
-#| echo: false
-#| warning: false
-#| message: false
-
-theft_recovery <- auto_Data |>
-  select(neighbourhood) |> 
-  count(neighbourhood)
-
-library(ggplot2)
-
-# Assuming 'years_vehicle' is your data frame
-
-years_vehicle <- theft_recovery |>
-  mutate(vehicle_year = as.numeric(neighbourhood))  # Convert to numeric if 'vehicle_year' is factor
-
-# Create a scatter plot
-ggplot(years_vehicle, aes(x = neighbourhood, y = n, fill = neighbourhood)) +
-  geom_point() +  
-  theme_minimal() + 
-  theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")
-
-```
-
 # Discussion

### 2024-04-23 paper/paper.qmd
new updates

--- 2024-04-22 paper/paper.qmd
@@
 #| label: averageTheftYear
-#| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
+#| fig-cap: "Average Auto Theft Crimes Reported in ottawa from 2018-2023"
 #| echo: false
@@
 ```{r}
-#| label: theft-brands
-#| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
-#| fig-subcap: ["Illnesses","Visits to the doctor"]
-#| layout-ncol: 3
+#| label: theftBrands
+#| fig-cap: "Top 10 Car Brands Stolen from 2018-2023"
+#| fig-subcap: ["Top 10 Brands Affected in 2018", "Top 10 Brands Affected in 2019", "Top 10 Brands Affected in 2020", "Top 10 Brands Affected in 2021", "Top 10 Brands Affected in 2022", "Top 10 Brands Affected in 2023"]
+#| layout-ncol: 2
 #| echo: false
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2018")
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
 
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2019")
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
 
@@
                                                         "LEXS" = "Lexus",
-                                                        "MERZ" = "Mercedes-Benz ",
+                                                        "MERZ" = "Mercedes",
                                                         "MAZD" = "Mazda",
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2020")
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
 
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2021")
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
 
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
 
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2023")
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
 
@@
 #| label: theft-recovery
-#| fig-cap: "Recovery Rates from 2018 - 2023"
+#| fig-cap: "Recovery Rates from 2018-2023"
 #| echo: false
@@
 
+# Filter for Count of Cars Recovered from 2018-2023
 theft_recovery <- auto_Data |>
@@
 
+# Create a Stacked Bar Graph of the Data
 theft_recovery |>
@@
   theme(legend.position = "bottom")
+
 ```
+
 ```{r}
 #| label: theft-neighborhoods
-#| fig-cap: "Recovery Rates from 2018 - 2023"
+#| fig-cap: "Top 5 Neighborhoods with Most Auto Thefts from 2018-2023"
 #| echo: false
@@
 
+# Filter for top 10 neighborhoods with the highest auto theft rates from 2018-2023
 theft_recovery <- auto_Data |>
@@
   theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022") 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
+
+```
+
+```{r}
+#| label: vehicleYearStolen
+#| fig-cap: "Vehicle Years Stolen from 2018-2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Count for vehicle year 
+years_vehicle <- auto_Data |>
+  select(vehicle_year) |> 
+  count(vehicle_year)
+
+# Code from Chat GPT (see LLM)
+
+years_vehicle <- years_vehicle |>
+  mutate(vehicle_year = as.numeric(vehicle_year))  # Convert to numeric if 'vehicle_year' is factor
+
+# Create a scatter plot
+ggplot(years_vehicle, aes(x = vehicle_year, y = n, fill = n)) +
+  geom_point() +  # Use geom_point() for scatter plot
+  scale_x_continuous(limits = c(1930, 2024)) +  # Adjust the x-axis limits here
+  theme_minimal() + 
+  theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
+
 ```

### 2024-04-24 paper/paper.qmd
updated paper, llm

@@
 title: "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
+sub-title: "An Analysis on top car brands stolen, recovery rate, and top neighborhoods with high theft rates"
 author: 
@@
 abstract: "The surge in auto theft rates across Canada has prompted car owners to become more vigilant regarding their vehicle's security.
-This paper gives a thorough examination of auto theft occurrences within Canada's capital, Ottawa.
-Focusing on the years 2018-2022, the paper provides a deep dive into the types of vehicles brands, model, and color targeted and offers insights into the neighborhoods within Ottawa exhibiting elevated rates of auto theft.
-Additionally, the study assesses factors such as the recovery rate of stolen cars, and provides solutions to improve vehicle security by understanding auto theft dynamics in the region."
+This paper examines auto theft occurrences within Canada's capital, Ottawa, Ontario.
+Focusing on the years 2018-2023, the paper considers the types of vehicles, brands, and neighborhoods targeted within Ottawa.
+Additionally, the study assesses factors such as the recovery rate of stolen cars, and compares the data to other regions of the country."
 format: pdf
@@
 #| include: false
+#| echo: false
 #| warning: false
@@
 library(kableExtra)
-library(maps)
 library(ggmap)
 library(arrow)
+library(knitr)
+library(tinytex)
+library(readr)
+library(janitor)
+library(dplyr)
 
@@
 
-In recent years, there is a significant increase of auto theft rates across various cities in Canada.
+There is no doubt that auto theft has become an increasingly popular topic over the last few years.
+Canada has seen exceeding rates of auto theft, where many thefts are starting to become more frequent and more violent.
+Between the years 2021-2023, Ontario has seen a 48.2% increase, while Quebec has seen a 57.9% increase [@citeEquite].
+Major cities like Toronto reported to have more than 12,000 vehicles stolen in 2023, which is equal to having one car stolen every 40 minutes [@citeCBC].
+In addition, a report from Canadian Finance & Leasing Association (CFLA) stated that Toronto saw a 300 per cent increase in vehicle thefts from 2015 to 2022  [@citeNational].
+
 The country’s capital, Ottawa has reported 221 stolen vehicles cases since January 1, 2024 [@citeCTV].
@@
 
-[TO BE EDITED]
-
-This paper aims to identify the gravity of auto theft as a pressing concern, particularly in Canada, in the @Introduction.
+This paper aims to provide a deep analysis of auto theft rates in Ottawa, Canada.
 Subsequently, the @Data section focuses on the methodology for data collection and data cleaning.
@@
 
-The data was cleaned by using the open source statistical programming language R [@citeR]...[TO UPDATE]
+The data was cleaned by using the open source statistical programming language R [@r], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `here` [@here], `kableExtra` [@kableExtra], `janitor` [@janitor], and `knitr` [@knitr].
+
+## Data Cleaning
+
+For the data cleaning process, many variables from the original dataset was removed to focus on the necessary variables.
+From the original dataset, the following variables were selected: vehicle_make, vehicle_model, vehicle_year, vehicle_colour, vehicle_value, occurred_date, reported_date, weekday, recovered, year, neighborhood.
+For clarity purposes, the variable “year” was renamed to “theft_year” to make it clear that the variable accounted for the year the vehicle was stolen.
 
@@
 
+## Average Count of Auto Theft Cases from 2018-2023 in Ottawa
+
 ```{r}
-#| label: averageTheftYear
-#| fig-cap: "Average Auto Theft Crimes Reported in ottawa from 2018-2023"
+#| label: fig-averageTheftYear
+#| fig-cap: "Average Auto Theft Crimes Reported in Ottawa from 2018-2023"
 #| echo: false
@@
 ```
+[@fig-averageTheftYear] shows the average number of auto theft cases reported in Ottawa from 2018 to 2023.
+2022 stands out with the highest number of cases, reaching a record high of 2,159.
+Conversely, 2023 saw the lowest number of cases, totaling 693.
+In 2018, 914 cases were reported, followed by 1012 in 2019.
+The number of reported auto theft cases decreased to 891 in 2020, but then surged to 1157 in 2021.
+
+## Top Brands Stolen from 2018-2023
+
+[@fig-theftBrands] presents six graphical representations showcasing the top 10 brands with the highest number of auto theft cases spanning the years 2018 to 2023.
+Across these six graphs, Toyota consistently emerges with the highest average of stolen cars, except for 2022, where Honda takes the lead in the number of cars stolen.
+A notable trend observed is the declining numbers for Ford, with a steady decrease in stolen cars over the years.
+Throughout all six graphs, the most frequently occurring brands include Toyota, Honda, Chevrolet, Dodge, Ford, Hyundai, and Nissan.
+
+[FIGURE 2.5] illustrates a scatter plot depicting the model year of stolen cars in Ottawa.
+The graph reveals that cars manufactured between 2000 and 2024 are predominantly targeted.
+Particularly, cars from the year 2021 are the most commonly stolen.
+
 ```{r}
-#| label: theftBrands
+#| label: fig-theftBrands
 #| fig-cap: "Top 10 Car Brands Stolen from 2018-2023"
@@
 ```{r}
-#| label: theft-recovery
-#| fig-cap: "Recovery Rates from 2018-2023"
-#| echo: false
-#| warning: false
-#| message: false
-
-# Filter for Count of Cars Recovered from 2018-2023
-theft_recovery <- auto_Data |>
-  select(theft_year, recovered)
-
-# Create a Stacked Bar Graph of the Data
-theft_recovery |>
-  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
-  geom_bar(position = "dodge2") +
-  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
-  theme(legend.position = "bottom")
-
-```
-
-```{r}
-#| label: theft-neighborhoods
-#| fig-cap: "Top 5 Neighborhoods with Most Auto Thefts from 2018-2023"
-#| echo: false
-#| warning: false
-#| message: false
-
-# Filter for top 10 neighborhoods with the highest auto theft rates from 2018-2023
-theft_recovery <- auto_Data |>
-  select(neighbourhood) |> 
-  count(neighbourhood) |> 
-  filter(n > 180)
-
-theft_recovery |>
-  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
-  geom_bar(stat = "identity") +
-  theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
-
-```
-
-```{r}
-#| label: vehicleYearStolen
+#| label: fig-vehicleYearStolen
 #| fig-cap: "Vehicle Years Stolen from 2018-2023"
@@
 
+## Recovery Rate
+
+@fig-theftrecovery illustrates the number of cars recovered and unrecovered from 2018 to 2023.
+Overall, it is evident that more cars are recovered, except for 2023, where more cars were left unrecovered compared to those recovered.
+From 2018 to 2021, the number of cars recovered consistently surpassed those left unrecovered.
+Specifically, in 2018, 570 cars were recovered out of 914 (as shown in Figure 1), reflecting a 62% recovery rate.
+Similarly, in 2019, with a total of 1,012 cars stolen (as shown in Figure 1), 646 were recovered (as shown in Figure 3), maintaining a recovery rate of 63%.
+
+In 2020, 529 out of 891 stolen cars were recovered, resulting in a 59% recovery rate.
+The following year, 2021, witnessed a slightly lower recovery rate of 53%, with 623 cars recovered out of 1,157 stolen.
+Notably, 2022 marked the highest number of both recovered and unrecovered cars, totaling 1,080 and 1,079 respectively.
+As per [FIGURE 1], a total of 2,159 cars were stolen in Ottawa in 2022.
+[FIGURE 3] indicates that 1,080 out of 2,159 were recovered, equating to a 50% recovery rate for 2022.
+
+However, 2023 stands out as the only year where the number of cars recovered is fewer than those left unrecovered.
+In that year, a total of 693 cars were stolen (as shown in Figure 1), and only 334 cars were recovered (as shown in Figure 3), resulting in a recovery rate of 48%.
+[FIGURE 3] illustrates that between 2018 and 2023, the recovery rate averages roughly between 60% and 50%.
+
+```{r}
+#| label: fig-theftrecovery
+#| fig-cap: "Recovery Rates from 2018-2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for Count of Cars Recovered from 2018-2023
+theft_recovery <- auto_Data |>
+  select(theft_year, recovered)
+
+# Create a Stacked Bar Graph of the Data
+theft_recovery |>
+  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
+  geom_bar(position = "dodge2") + 
+  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
+  theme(legend.position = "bottom")
+
+```
+
+## Top Neighborhoods
+
+@tbl-theftneighborhoods2018 displays the top 5 neighborhoods with the highest rates of auto theft in Ottawa in 2018.
+Among these, Overbrook – McArthur recorded the highest number of cases, totaling 46.
+@tbl-theftneighborhoods2023 presents the top 5 neighborhoods with the highest rates of auto theft in 2023, with Portobello South leading with 33 reported cases.
+Notably, Overbrook – McArthur maintains its position in the top 5, reporting 21 cases of auto theft.
+@tbl-theftneighborhoods showcases the top 5 neighborhoods with the highest rates of auto theft cumulatively from 2018 to 2023.
+In this comprehensive view, Centretown emerges with the highest number of auto theft cases, totaling 298.
+Overbrook – McArthur follows closely with 285 cases, while Portobello South and Stittsville tie with 218 cases each, and East Industrial reports a total of 182 cases.
+
+```{r}
+#| label: tbl-theftneighborhoods2018
+#| tbl-cap: "Table of Top 5 Neighborhoods with Most Auto Thefts in 2018"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for top 10 neighborhoods with the highest auto theft rates from 2018
+theft_neighbourhood2018 <- auto_Data |>
+  filter(theft_year == "2018") |>
+  select(neighbourhood) |> 
+  count(neighbourhood) |> 
+  filter(n > 20)
+
+theft_neighbourhood2018 |>
+  kable(
+    col.names = c("Neighborhood", "Total Number of Cases") , 
+    align = c("c", "c")
+  )
+
+```
+
+```{r}
+#| label: tbl-theftneighborhoods2023
+#| tbl-cap: "Table of Top 5 Neighborhoods with Most Auto Thefts in 2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for top 10 neighborhoods with the highest auto theft rates from 2023
+theft_neighbourhood2023 <- auto_Data |>
+  filter(theft_year == "2023") |>
+  select(neighbourhood) |> 
+  count(neighbourhood) |> 
+  filter(n > 20)
+
+theft_neighbourhood2023 |>
+  kable(
+    col.names = c("Neighborhood", "Total Number of Cases") , 
+    align = c("c", "c")
+  )
+
+```
+
+```{r}
+#| label: tbl-theftneighborhoods
+#| tbl-cap: "Table of Top 5 Neighborhoods with Most Auto Thefts from 2018-2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for top 10 neighborhoods with the highest auto theft rates from 2018-2023
+theft_neighbourhood <- auto_Data |>
+  select(neighbourhood) |> 
+  count(neighbourhood) |> 
+  filter(n > 180)
+
+theft_neighbourhood |>
+  kable(
+    col.names = c("Neighborhood", "Total Number of Cases") , 
+    align = c("c", "c")
+  )
+```
+
 # Discussion

### 2024-04-24 paper/paper.qmd
updated paper

@@
 title: "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
-sub-title: "An Analysis on top car brands stolen, recovery rate, and top neighborhoods with high theft rates"
+subtitle: "An Analysis on top car brands stolen, recovery rate, and top neighborhoods with high theft rates"
 author: 
@@
 This paper examines auto theft occurrences within Canada's capital, Ottawa, Ontario.
-Focusing on the years 2018-2023, the paper considers the types of vehicles, brands, and neighborhoods targeted within Ottawa.
-Additionally, the study assesses factors such as the recovery rate of stolen cars, and compares the data to other regions of the country."
+Focusing on the years 2018-2023, the paper considers the types of vehicles, brands, and neighborhoods targeted within Ottawa as well as the recovery rate.
+Additionally, the study aims to discuss reasons for auto theft and comparisons to data in province."
 format: pdf
@@
 This paper aims to provide a deep analysis of auto theft rates in Ottawa, Canada.
+The estimand for this paper is to quantify the impact of various factors on the occurrence and frequency of auto thefts in Ottawa, ultimately aiming to understand the dynamics driving the escalating trend in auto theft rates within the city.
 Subsequently, the @Data section focuses on the methodology for data collection and data cleaning.
 The @Results section delves into a comprehensive analysis of auto theft patterns in Ottawa from the years 2018-2022.
-Finally, in the @Discussion section, the paper explores potential drivers behind the escalation in auto theft rates, strategies for enhancing vehicle security, and prospective trends in auto theft.
+Finally, in the @Discussion section, the paper explores potential drivers behind the escalation in auto theft rates and comparing the data to trends across Ontario.
 
@@
 #| label: fig-averageTheftYear
-#| fig-cap: "Average Auto Theft Crimes Reported in Ottawa from 2018-2023"
+#| fig-cap: "Average Auto Theft Crimes Reported in ottawa from 2018-2023"
 #| echo: false
@@
 
-[@fig-theftBrands] presents six graphical representations showcasing the top 10 brands with the highest number of auto theft cases spanning the years 2018 to 2023.
+@fig-theftBrands presents six graphical representations showcasing the top 10 brands with the highest number of auto theft cases spanning the years 2018 to 2023.
 Across these six graphs, Toyota consistently emerges with the highest average of stolen cars, except for 2022, where Honda takes the lead in the number of cars stolen.
@@
 
-[FIGURE 2.5] illustrates a scatter plot depicting the model year of stolen cars in Ottawa.
+@fig-vehicleYearStolen illustrates a scatter plot depicting the model year of stolen cars in Ottawa.
 The graph reveals that cars manufactured between 2000 and 2024 are predominantly targeted.
@@
 ```{r}
+#| label: fig-theft-recovery
+#| fig-cap: "Recovery Rates from 2018-2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for Count of Cars Recovered from 2018-2023
+theft_recovery <- auto_Data |>
+  select(theft_year, recovered)
+
+# Create a Stacked Bar Graph of the Data
+theft_recovery |>
+  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
+  geom_bar(position = "dodge2") +
+  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+#| label: fig-theft-neighborhoods
+#| fig-cap: "Top 5 Neighborhoods with Most Auto Thefts from 2018-2023"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for top 10 neighborhoods with the highest auto theft rates from 2018-2023
+theft_recovery <- auto_Data |>
+  select(neighbourhood) |> 
+  count(neighbourhood) |> 
+  filter(n > 180)
+
+theft_recovery |>
+  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
+  geom_bar(stat = "identity") +
+  theme_minimal() + theme(legend.position = "bottom") + 
+  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
+
+```
+
+```{r}
 #| label: fig-vehicleYearStolen
@@
 
-## Methods to Increase Vehicle Security
+It is evident that auto theft reports are increasing.
+However, why are cases increasing and what happens to cars that are stolen?
+From @fig-vehicleYearStolen, we understand that vehicles made between 2000-2024, are most targeted for theft.
+Cars made in 2021 also had the highest rate of theft (@fig-vehicleYearStolen).
+A report by Equite, a not-for-profit organization that focuses on reducing and preventing insurance fraud and crime in Canada, states that the majority of stolen vehicles are new as organized crime rings focus their efforts on stealing new and luxury vehicles for maximum profit in overseas sales [@citeEquite2].
+Most newer vehicles also have a push button start, which makes it easier for criminals to steal.
+In fact, some criminals will target certain makes/models as they are already equipped to steal, either through hardware or software hacking [@citeEquite2].
+Criminals will also look for vehicles without anti-theft protection (video/audio devices, steering wheel locks, etc) as it is easier to steal.
 
-## Auto Theft Rates Projection
+## Comparison to Ontario
 
-## Weaknesses and next steps
+Although this paper focuses specifically on auto theft rates in Ottawa, how different are the trends compared to Ontario?
+According to a report by Equite, the majority of vehicle thefts in Ontario were vehicles made in 2017 or newer [@citeEquite].
+This can be seen in [FIGURE] where newer vehicles made in 2017 or newer had a higher rate of theft in Ottawa.
+
+## Weaknesses and Next Steps
+
+Some limitations of this paper include the reliability of the dataset.
+While the data originates from the Ottawa Police and relies on reports submitted by Ottawa residents, the accuracy of these reports may vary.
+The dataset excludes instances of stolen vehicle cases that went unreported, potentially skewing the overall picture of vehicle theft within the community.
+This absence of unreported cases could lead to an underestimation of the true extent of the issue, limiting the comprehensiveness of the findings presented in this paper.
+
+# Conclusion
+
+In conclusion, the surge in auto theft incidents, both in Ottawa and across Canada, underscores a pressing issue that demands attention.
+The escalation in theft rates poses a significant threat to public safety, prompting concerns about the effectiveness of current preventative measures and law enforcement efforts.
+The analysis conducted in this paper sheds light on the auto theft landscape specifically in Ottawa, offering valuable insights into the trends and patterns observed between 2018 and 2023.
+By examining variables such as the types of vehicles targeted, the recovery rates, and the geographical distribution of thefts across different neighborhoods, this research provides a comprehensive understanding of the dynamics driving the escalating trend in auto theft.
+Moving forward, addressing the root causes behind the surge in auto theft requires a multifaceted approach that combines effective law enforcement strategies, community engagement initiatives, and technological solutions to deter criminals and protect vulnerable vehicles.
+By leveraging the findings of this study and collaborating with stakeholders, policymakers, and law enforcement agencies can develop targeted interventions to curb the rising tide of auto theft and safeguard the well-being of Ottawa residents.
 

### 2024-04-25 paper/paper.qmd
updated paper

@@
 title: "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
-subtitle: "An Analysis on top car brands stolen, recovery rate, and top neighborhoods with high theft rates"
+subtitle: "An Analysis on Top Car Brands Stolen, Recovery Rate, and Top Neighborhoods with High Theft Rates"
 author: 
@@
 library(kableExtra)
-library(ggmap)
 library(arrow)
@@
 library(dplyr)
+library(ggplot2)
 
@@
 
-# Introduction
+# Introduction {#sec-intro}
 
@@
 The estimand for this paper is to quantify the impact of various factors on the occurrence and frequency of auto thefts in Ottawa, ultimately aiming to understand the dynamics driving the escalating trend in auto theft rates within the city.
-Subsequently, the @Data section focuses on the methodology for data collection and data cleaning.
-The @Results section delves into a comprehensive analysis of auto theft patterns in Ottawa from the years 2018-2022.
-Finally, in the @Discussion section, the paper explores potential drivers behind the escalation in auto theft rates and comparing the data to trends across Ontario.
+Subsequently, the @sec-data section focuses on the methodology for data collection and data cleaning.
+The @sec-Results section delves into a comprehensive analysis of auto theft patterns in Ottawa from the years 2018-2022.
+Finally, in the @sec-Discussion section, the paper explores potential drivers behind the escalation in auto theft rates and comparing the data to trends across Ontario.
 
@@
 
-The data was cleaned by using the open source statistical programming language R [@r], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `here` [@here], `kableExtra` [@kableExtra], `janitor` [@janitor], and `knitr` [@knitr].
+The dataset used for this paper was provided by Ottawa Police [@citeOttawa].
+The dataset contains theft of motor vehicle occurrencesfrom the years 2018-2023.The data was cleaned by using the open source statistical programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `kableExtra` [@kableExtra], `janitor` [@janitor], arrow [@arrow] ,and `knitr` [@knitr].
 
@@
 
-# Results
+# Results {#sec-Results}
 
@@
 ```{r}
-#| label: fig-theft-recovery
-#| fig-cap: "Recovery Rates from 2018-2023"
-#| echo: false
-#| warning: false
-#| message: false
-
-# Filter for Count of Cars Recovered from 2018-2023
-theft_recovery <- auto_Data |>
-  select(theft_year, recovered)
-
-# Create a Stacked Bar Graph of the Data
-theft_recovery |>
-  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
-  geom_bar(position = "dodge2") +
-  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
-  theme(legend.position = "bottom")
-
-```
-
-```{r}
-#| label: fig-theft-neighborhoods
-#| fig-cap: "Top 5 Neighborhoods with Most Auto Thefts from 2018-2023"
-#| echo: false
-#| warning: false
-#| message: false
-
-# Filter for top 10 neighborhoods with the highest auto theft rates from 2018-2023
-theft_recovery <- auto_Data |>
-  select(neighbourhood) |> 
-  count(neighbourhood) |> 
-  filter(n > 180)
-
-theft_recovery |>
-  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
-  geom_bar(stat = "identity") +
-  theme_minimal() + theme(legend.position = "bottom") + 
-  labs(x = "Vehicle Brand", y = "Count", fill = "Count")
-
-```
-
-```{r}
 #| label: fig-vehicleYearStolen
@@
 From 2018 to 2021, the number of cars recovered consistently surpassed those left unrecovered.
-Specifically, in 2018, 570 cars were recovered out of 914 (as shown in Figure 1), reflecting a 62% recovery rate.
-Similarly, in 2019, with a total of 1,012 cars stolen (as shown in Figure 1), 646 were recovered (as shown in Figure 3), maintaining a recovery rate of 63%.
+Specifically, in 2018, 570 cars were recovered out of 914 (@fig-averageTheftYear), reflecting a 62% recovery rate.
+Similarly, in 2019, with a total of 1,012 cars stolen (@fig-averageTheftYear), 646 were recovered (@fig-theftrecovery), maintaining a recovery rate of 63%.
 
@@
 Notably, 2022 marked the highest number of both recovered and unrecovered cars, totaling 1,080 and 1,079 respectively.
-As per [FIGURE 1], a total of 2,159 cars were stolen in Ottawa in 2022.
-[FIGURE 3] indicates that 1,080 out of 2,159 were recovered, equating to a 50% recovery rate for 2022.
+As per @fig-averageTheftYear, a total of 2,159 cars were stolen in Ottawa in 2022.
+@fig-theftrecovery indicates that 1,080 out of 2,159 were recovered, equating to a 50% recovery rate for 2022.
 
-However, 2023 stands out as the only year where the number of cars recovered is fewer than those left unrecovered.
-In that year, a total of 693 cars were stolen (as shown in Figure 1), and only 334 cars were recovered (as shown in Figure 3), resulting in a recovery rate of 48%.
-[FIGURE 3] illustrates that between 2018 and 2023, the recovery rate averages roughly between 60% and 50%.
+However, 2023 stands out as the only year where the number of cars recovered is fewer than those left unrecoverable.
+A total of 693 cars were stolen (as shown in @fig-averageTheftYear), and only 334 cars were recovered (as shown in @fig-theftrecovery), resulting in a recovery rate of 48%.
+@fig-theftrecovery illustrates that between 2018 and 2023, the recovery rate averages roughly between 60% and 50%.
 
@@
 ## Top Neighborhoods
-
-@tbl-theftneighborhoods2018 displays the top 5 neighborhoods with the highest rates of auto theft in Ottawa in 2018.
-Among these, Overbrook – McArthur recorded the highest number of cases, totaling 46.
-@tbl-theftneighborhoods2023 presents the top 5 neighborhoods with the highest rates of auto theft in 2023, with Portobello South leading with 33 reported cases.
-Notably, Overbrook – McArthur maintains its position in the top 5, reporting 21 cases of auto theft.
-@tbl-theftneighborhoods showcases the top 5 neighborhoods with the highest rates of auto theft cumulatively from 2018 to 2023.
-In this comprehensive view, Centretown emerges with the highest number of auto theft cases, totaling 298.
-Overbrook – McArthur follows closely with 285 cases, while Portobello South and Stittsville tie with 218 cases each, and East Industrial reports a total of 182 cases.
 
@@
 
-# Discussion
+@tbl-theftneighborhoods2018 displays the top 5 neighborhoods with the highest rates of auto theft in Ottawa in 2018.
+Among these, Overbrook – McArthur recorded the highest number of cases, totaling 46.
+@tbl-theftneighborhoods2023 presents the top 5 neighborhoods with the highest rates of auto theft in 2023, with Portobello South leading with 33 reported cases.
+Notably, Overbrook – McArthur maintains its position in the top 5, reporting 21 cases of auto theft.
+@tbl-theftneighborhoods showcases the top 5 neighborhoods with the highest rates of auto theft cumulatively from 2018 to 2023.
+In this comprehensive view, Centretown emerges with the highest number of auto theft cases, totaling 298.
+Overbrook – McArthur follows closely with 285 cases, while Portobello South and Stittsville tie with 218 cases each, and East Industrial reports a total of 182 cases.
+
+# Discussion {#sec-Discussion}
 
@@
 According to a report by Equite, the majority of vehicle thefts in Ontario were vehicles made in 2017 or newer [@citeEquite].
-This can be seen in [FIGURE] where newer vehicles made in 2017 or newer had a higher rate of theft in Ottawa.
+This can be seen in @fig-vehicleYearStolen where newer vehicles made in 2017 or newer had a higher rate of theft in Ottawa.
 
@@
 
-In conclusion, the surge in auto theft incidents, both in Ottawa and across Canada, underscores a pressing issue that demands attention.
+In conclusion, the surge in auto theft incidents, both in Ottawa and across Canada, is a pressing issue that demands attention.
 The escalation in theft rates poses a significant threat to public safety, prompting concerns about the effectiveness of current preventative measures and law enforcement efforts.
-The analysis conducted in this paper sheds light on the auto theft landscape specifically in Ottawa, offering valuable insights into the trends and patterns observed between 2018 and 2023.
+The analysis conducted in this paper sheds light on the auto theft landscape specifically in Ottawa, offering insights into the trends and patterns observed between 2018 and 2023.
 By examining variables such as the types of vehicles targeted, the recovery rates, and the geographical distribution of thefts across different neighborhoods, this research provides a comprehensive understanding of the dynamics driving the escalating trend in auto theft.
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-\newpage
-
 # References

### 2024-04-25 paper/paper.qmd
updated paper, references

@@
 
+The paper uses the data provided by Ottawa Police [@citeOttawa].
+The dataset defines auto theft as the act of taking or another person's vehicle [@citeOttawa].The data was cleaned by using the open source statistical programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `kableExtra` [@kableExtra], `janitor` [@janitor], arrow [@arrow] ,and `knitr` [@knitr].
+
 # Data {#sec-data}
 
-The dataset used for this paper was provided by Ottawa Police [@citeOttawa].
-The dataset contains theft of motor vehicle occurrencesfrom the years 2018-2023.The data was cleaned by using the open source statistical programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `kableExtra` [@kableExtra], `janitor` [@janitor], arrow [@arrow] ,and `knitr` [@knitr].
+The dataset provided by Ottawa Police contains all theft of motor vehicle occurrences from the years 2018-2023. The data is updated annually, and is available for public use.
 

### 2024-04-25 paper/paper.qmd
updated paper, scripts

@@
 For the data cleaning process, many variables from the original dataset was removed to focus on the necessary variables.
-From the original dataset, the following variables were selected: vehicle_make, vehicle_model, vehicle_year, vehicle_colour, vehicle_value, occurred_date, reported_date, weekday, recovered, year, neighborhood.
+From the original dataset, the following variables were selected: vehicle_make, vehicle_model, vehicle_year, recovered, year, neighborhood.
 For clarity purposes, the variable “year” was renamed to “theft_year” to make it clear that the variable accounted for the year the vehicle was stolen.

### 2024-04-25 scripts/01-download_data.R

@@ -12,14 +12,10 @@ library(tidyverse)
 
 #### Download data ####
 
-# code from CHAT GPT
-# Set the file path to your CSV file
+# Go to https://data.ottawapolice.ca/datasets/e5453a203e2a4baeac32ac4e70ce852c_0/about
 
-# Read the CSV file
-auto_theft_data <- read.csv("data/analysis_data/Auto_Theft_Ottawa_Data.csv")
+# Click "Download" 
 
-#### Save data ####
-# saving original dataset in raw_data
-write.csv(auto_theft_data, "data/raw_data/Auto_Theft_Ottawa_Data.csv") 
+# Click "Download CSV"
 
-         
+# Manually add CSV to R Project

### 2024-04-25 scripts/02-data_cleaning.R

@@ -1,11 +1,9 @@
 #### Preamble ####
 # Purpose: Cleans the raw auto theft data collected by Ottawa Police
 # Author: Catherine Punnoose
-# Date: 28 March 2024
-# Contact: catherine.punnoose@gmail.com
+# Date: 24 April 2024
+# Contact: catherine.punnoose@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(dplyr)
@@ -20,9 +18,7 @@ raw_data <- read.csv("data/raw_data/Auto_Theft_Ottawa_Data.csv")
 cleaned_data <-
   raw_data |>
   janitor::clean_names() |>
-  select(vehicle_make, vehicle_model, vehicle_year, vehicle_colour, 
-         vehicle_value, occurred_date, reported_date, weekday, 
-         recovered, year, neighbourhood) |>
+  select(vehicle_make, vehicle_model, vehicle_year, recovered, year, neighbourhood) |>
   mutate(theft_year = year) # changing year to theft_year for clarity
 
 

### 2024-04-25 scripts/04-model.R

@@ -1,38 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(arrow)
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-analysis_data <- read_parquet("data/analysis_data/Auto_Theft_Ottawa_Data.csv")
-
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width, #sub with my own variables
-    data = analysis_data,
-    family = gaussian(),# spec for building a model
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
-
-

### 2024-04-25 paper/datasheet_template.qmd
updated datasheet

---
title: "Datasheet for 'Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa'"
subtitle: "An Analysis on Top Car Brands Stolen, Recovery Rate, and Top Neighborhoods with High Theft Rates"
author: Catherine Punnoose
thanks: "Code and data are available at: https://github.com/catherinee1216/AutoTheft_Ottawa.git."
date: today
date-format: long
abstract: "The surge in auto theft rates across Canada has prompted car owners to become more vigilant regarding their vehicle's security. This paper examines auto theft occurrences within Canada's capital, Ottawa, Ontario. Focusing on the years 2018-2023, the paper considers the types of vehicles, brands, and neighborhoods targeted within Ottawa as well as the recovery rate. Additionally, the study aims to discuss reasons for auto theft and comparisons to data in province."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Auto Theft reports in Ottawa, Canada. The dataset did not provide an explaination of a specific gap that needed to be filled. 
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by Ottawa Police.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - Ottawa Police. 
4. *Any other comments?*
    - N/A

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The dataset is compromised in the form of an excel spreadsheet. 
2. *How many instances are there in total (of each type, if appropriate)?*
	- None.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- N/A
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- The data consists of unprocessed text. 
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- N/A
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some rows in vehicle_value were empty as the investigators may not have been able to collect the value of the stolen vehicle. 
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- N/A
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- N/A
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- N/A
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained.  
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No. All data provided in the dataset is open for the public. 
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No. 
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No. 
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No. 
16. *Any other comments?*
	- No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was reports given by Ottawa residents to Ottawa Police. It can be assumed that each report was verified by the Police before adding it to the dataset. 
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- N/A
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- N/A
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Any Ottawa resident who experienced auto theft was involved in the data collection process. It is not confirmed if the residents were given any compensation. 
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Data was collected from 2018, and is continued to be collected today. 
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- N/A
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was collected by Ottawa Police. 
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- N/A
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- N/A
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- N/A
12. *Any other comments?*
	- N/A

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- N/A
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes. The raw data can be found at data/raw_data/Auto_Theft_Ottawa_Data.csv. 
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- No software was used. 
4. *Any other comments?*
	- N/A

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- It is unknown if the dataset has been used before. 
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- N/A
3. *What (other) tasks could the dataset be used for?*
	- The dataset can be used to determine specific details of auto theft cases such as top vehicle color stolen, comparing theft rates to other provinces or countires, etc. 
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for any task that is not related to auto theft and Ottawa. 
6. *Any other comments?*
	- N/A

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset can be accessed by any third party at https://data.ottawapolice.ca/datasets/e5453a203e2a4baeac32ac4e70ce852c_0/about
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is distributed through website. 
3. *When will the dataset be distributed?*
	- It is available on https://data.ottawapolice.ca/datasets/e5453a203e2a4baeac32ac4e70ce852c_0/about. 
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- N/A
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- N/A
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- N/A
7. *Any other comments?*
	- N/A

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Ottawa Police Services. 
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- N/A
3. *Is there an erratum? If so, please provide a link or other access point.*
	- N/A
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset is updated annually. 
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- N/A
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- N/A
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- N/A
8. *Any other comments?*
	- N/A

\newpage

# References

### 2024-04-25 paper/paper.qmd
updated READMe, tests, and paper

@@
                              ))
+# Create a Bar Graph
 brands2018_top10 |>
@@
                                                      ))
+# Create a Bar Graph
 brandtheft2019_top10 |>
@@
                                                      ))
+# Create a Bar Graph
 brandtheft2020_top10 |>
@@
                                                      ))
+# Create a Bar Graph
 brandtheft2021_top10 |>
@@
                                                      ))
+# Create a Bar Graph
 brandtheft2022_top10 |>
@@
 
-# filter for Top 10 Brands 
+# Filter for Top 10 Brands 
 brandtheft2023_top10 <- brand_theft_2023 |>
@@
                                                      ))
+# Create a Bar Graph
 brandtheft2023_top10 |>
@@
 
-# Create a Stacked Bar Graph of the Data
+# Create Stacked Bar Graph
 theft_recovery |>

### 2024-04-25 scripts/00-simulate_data.R

@@ -1,19 +1,76 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
+# Purpose: Simulates dataset of "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
 # Author: Catherine Punnoose
-# Date: 28 March 2024
+# Date: 24 April 2024
 # Contact: catherine.punnoose@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(janitor)
+library(ggplot2)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# Code Generated by Chat GPT 3.5 (see usage.txt)
+
+# Simulated Data Generation of Average Auto Theft Crimes Reported in ottawa from 2018-2023 Graph
+set.seed(123)
+auto_Data <- tibble(
+  theft_year = sample(2018:2023, 1000, replace = TRUE)
+)
+
+# Plotting
+auto_Data %>%
+  ggplot(aes(x = theft_year)) + 
+  labs(x = "Year", y = "Count") + 
+  theme_minimal() +
+  geom_histogram(stat = "count")
+
+
+# Simulated Data Generation of Top 10 Car Brands Stolen from 2018-2023 Graphs
+set.seed(123)
+auto_Data <- tibble(
+  theft_year = sample(2018:2023, 1000, replace = TRUE),
+  vehicle_make = sample(c("CHEV", "DODG", "FORD", "GMC", "HOND", "HYUN", "LEXS", "MAZD", "NISS", "TOYT"), 1000, replace = TRUE)
+)
+
+# Define a function to create the plot for each year
+create_brand_theft_plot <- function(year, data) {
+  brand_theft <- data %>%
+    filter(theft_year == year) %>%
+    count(vehicle_make) %>%
+    mutate(vehicle_make = recode(vehicle_make,
+                                 "CHEV" = "Chevrolet",
+                                 "DODG" = "Dodge",
+                                 "FORD" = "Ford",
+                                 "GMC" = "GMC",
+                                 "HOND" = "Honda",
+                                 "HYUN" = "Hyundai",
+                                 "LEXS" = "Lexus",
+                                 "MAZD" = "Mazda",
[…]

### 2024-04-25 scripts/03-test_data.R

@@ -1,17 +1,55 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests for the dataset of "Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa"
+# Author: Catherine Punnoose
+# Date: 24 April 2024
+# Contact: catherine.punnoose@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: reload 02-data_cleaning.R
 
 
 #### Workspace setup ####
 library(arrow)
 library(tidyverse)
+library(testthat)
+library(ggplot2)
 
-# use testthat package
+#### Reload Cleaned Data ####
+
+auto_Data <- read_parquet(here::here("data/analysis_data/analysis_data.parquet"))
 
 #### Test data ####
+
+# Code generated by Chat GPT 3.5 (See usage.txt.)
+# Testing Average Theft Year Histogram
+plot_theft_year_histogram <- function(data) {
+  data %>%
+    ggplot(aes(x = theft_year)) + 
+    labs(x = "Year", y = "Count") + 
+    theme_minimal() +
+    geom_histogram(stat = "count")
+}
+
+# Write test cases
+test_that("Theft year histogram plot is generated correctly", {
+  # Simulated Data Generation
+  set.seed(123)
+  auto_Data <- tibble(
+    theft_year = sample(2018:2023, 1000, replace = TRUE)
+  )
+  
+  # Generate the plot
+  plot <- plot_theft_year_histogram(auto_Data)
+  
+  # Check if the plot is of class ggplot
+  expect_true("ggplot" %in% class(plot))
+  
+  # Check if the x-axis label is correct
+  expect_equal(ggplot2:::ggtitle_string(plot$labels$x), "Year")
+  
+  # Check if the y-axis label is correct
+  expect_equal(ggplot2:::ggtitle_string(plot$labels$y), "Count")
+  
+  # Check if the plot contains geom_histogram
+  expect_true("GeomHistogram" %in% class(plot$layers[[1]]$geom))
+})
+
